### 2015-11-22 mirror-descent.tex
Few iprovements to Mirror Descent chapter.

\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The
Regularized Leader} but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on
past loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an
arbitrary way. If $R_t$ is not differentiable\footnote{Note that this can
happen even when $R_t$ is a restriction of a differentiable function defined on
a superset of $K$.  If $K$ is bounded and closed, $R_t$ fails to be
differentiable at the boundary of $K$. If $K$ is a subset of an affine subspace
of a dimension smaller than the dimension of $V$, then $R_t$ fails to be
differentiable everywhere.}, to define Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v \rangle$, the
algorithm has to choose together with $R_t$ a subgradient map $\grad R_t:K \to
V$.  If $R_t$ is a restriction of a differentiable function $R'_t$, it is convenient
to define $\grad R_t(w) = \grad R'_t(w)$ for all $w \in K$. The following lemma
bounds instantaneous regret of \textsc{Mirror Descent}.

\begin{lemma}[Regret of Mirror Descent]
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} for any $u \in K$ satisfies
$$
\Regret_T(u) \le \sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
$$
\end{lemma}

Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
prominent example is projected gradient descent defined by $w_{t+1} = \Pi_K(w_t
- \eta_t \ell_t)$ where $\Pi_K$ is the Euclidean projection to $K$ and
$\eta_t>0$ is the step size.  Projected gradient descent is an instance of
\textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
A special case of projected gradient descent with step size $\eta_t = \frac{1}{\sqrt{t}}$
is Zinkevich's \textsc{Generalized Infinitesimal Gradient Ascent} (GIGA)
algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal versions of \textsc{AdaGrad with Composite Mirror
Descent Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of
\textsc{Mirror Descent} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2} \; .
$$
respectively. When $d=1$, both algorithms are equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm.
Also note that regularizer \eqref{equation:ada-grad} can be viewed
as a scale-free generalization of Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer} coincides with GIGA.

@@

\begin{theorem}[Regret of Scale-Free Mirror Descent]
\label{theorem:regret-scale-free-mirror-descent}
Suppose $K \subseteq V$ is a non-empty closed convex subset. Suppose
that $R:K \to \R$ is a $\lambda$-strongly convex function with respect to a norm $\|\cdot\|$.
\textsc{Scale-Free Mirror Descent} with regularizer $R$ satisfies
$$
\Regret_T(u) \le \left( \frac{1}{\lambda} + \sup_{v \in K} \Breg_R(u,v) \right) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{theorem}

We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function
and optimize $\lambda$.

@@

The bounds in Theorem~\ref{theorem:regret-scale-free-mirror-descent} and
Corollary~\ref{corollary:regret-scale-free-mirror-descent} are vacuous when
$\Breg_R(u,v)$ is not bounded. One might wonder if the
assumption that $\Breg_R(u,v)$ is bounded is necessary. We show necessity of
this assumption on two counter-examples where $\Breg_R(u,v)$ is not bounded.
In these counter-examples, we construct sequences of loss vectors $\ell_1,
\ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots =
\|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has $\Omega(T)$
or worse regret.

>>>>

\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The
Regularized Leader} but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, to define
Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
\rangle$, the algorithm has to choose together with $R_t$ a subgradient map
$\grad R_t:K \to V$.  If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.

\begin{lemma}[Regret of Mirror Descent]
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
$u \in K$,
$$
\Regret_T(u)
\le
\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
$$
\end{lemma}

Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
prominent example is projected gradient descent defined by $w_{t+1} = \Pi_K(w_t
- \eta_t \ell_t)$ where $\Pi_K$ is the Euclidean projection to $K$ and
$\eta_t>0$ is the step size.  Projected gradient descent is an instance of
\textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
A special case of projected gradient descent with step size $\eta_t =
\frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized Infinitesimal Gradient
Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal
versions of \textsc{AdaGrad with Composite Mirror Descent
Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{Mirror
Descent} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2} \; .
$$
respectively. When $d=1$, both algorithms are equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad} can be viewed as a scale-free generalization of
Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer}
coincides with GIGA.

@@

\begin{theorem}[Regret of Scale-Free Mirror Descent]
\label{theorem:regret-scale-free-mirror-descent}
Suppose $K \subseteq V$ is a non-empty closed convex subset. Suppose that $R:K
\to \R$ is a $\lambda$-strongly convex function with respect to a norm
$\|\cdot\|$.  \textsc{Scale-Free Mirror Descent} with regularizer $R$ satisfies
for any $u \in K$,
$$
\Regret_T(u) \le \left( \frac{1}{\lambda} + \sup_{v \in K} \Breg_R(u,v) \right) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{theorem}

We choose regularizer $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex
function and optimize $\lambda$. The result is the following Corollary. Its proof is trivial.

@@

The bounds in Theorem~\ref{theorem:regret-scale-free-mirror-descent} and
Corollary~\ref{corollary:regret-scale-free-mirror-descent} are vacuous when
$\Breg_R(u,v)$ is not bounded. One might wonder if the assumption that
$\Breg_R(u,v)$ is bounded is necessary. We show necessity of this assumption on
two counter-examples.  In these
counter-examples, we consider strongly convex regularizers $R$ such that
$\Breg_R(u,v)$ is not bounded and we construct sequences of loss vectors
$\ell_1, \ell_2, \dots, \ell_T$ such that $\|\ell_1\|_* = \|\ell_2\|_* = \dots
= \|\ell_T\|_* = 1$ and \textsc{Scale-Free Mirror Descent} has $\Omega(T)$ or
worse regret.
